Learning constraints in spreadsheets and tabular data

### 2016-05-27 Paper/ECAI/tabular.tex
Transferred all comments

%%\ecaisubmission   % inserts page numbers. Use only for submission of paper.
                  % Do NOT use for camera-ready version of paper.

@@

\title{Learning constraints in tabular data}
% see http://www.acm.org/binaries/content/assets/publications/article-templates/sig-alternate-sample.tex

@@

\begin{abstract}
%Spreadsheets, comma separated value files and other tabular data representations are in wide use today. However, the correct use and maintenance of functions in and between the tabular data can be error-prone and overwhelming. In this work, we investigate the automatic discovery of constraints (functions and relations) from raw tabular data. Our method takes inspiration form inductive logic programming and constraint satisfiability. We represent common spreadsheet functions as predicates with arguments that must satisfy number of constraints over the arguments. This allows generic constraint satisfaction techniques to be used to find all such satisfying predicates/functions within the rows, columns and submatrices of the tables. We show the effectiveness and accuracy of this approach on a number of spreadsheets from varying sources.
Spreadsheets, CSV (comma separated value) files and other tabular data representations are in wide use today.
However, modeling, maintaining and discovering formulas in tabular data and between spreadsheets can be time consuming and error-prone.
In this work, we investigate the automatic discovery of constraints (functions and relations) from raw tabular data.
We see multiple promising applications for this technique, e.g. in rediscovering constraints, auto-completion and error checking.
Our method takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
Common spreadsheet functions are represented as predicates whose arguments must satisfy a number of constraints.
Constraint learning techniques are used to identify predicates (constraints) that hold between blocks of rows and columns of the tables.
%New types of constraints can be easily added to the system by specifying them in a declarative language.
We show that our approach is able to accurately discover constraints in spreadsheets from various sources.
\end{abstract}

@@

The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where
the discovery of clausal constraints has been studied in an inductive logic programming setting \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.  It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from
very few examples.  Our approach contrasts with these in that it focusses
on learning both column- and row-constraints, as does Modelseeker~\cite{modelseeker} which is restricted to traditional CSPs. It contrasts with Flashfill in that it can learn from numeric data too, as well as general constraints in addition to functions.
A more detailed discussion on related work is contained in Section~\ref{sec:related_work}.

@@

\section{Formalization}\label{sec:formalization}
Our goal is to automatically discover constraints (functions and relations) between the rows and the columns of tables in a spreadsheet. This is applicable not just to data from spreadsheets, but any data in tabular form, hence the name.

@@

We first introduce some terminology and the concept of \template, after which we define the problem and make some additional considerations.

@@

A block has the following properties:
\begin{itemize}
\item \textit{type}: a block is type-consistent, so it has a type
\item \textit{table}: the table that the block belongs to
\item \textit{orientation}: either row-oriented or column-oriented
\item \textit{size}: the number of vectors a block contains
\item \textit{length}: the length of its vectors; as all vectors are from the same table, they always have the same length.
\item \textit{rows}: this is orientation specific, for row-based blocks this is the size, for column-based blocks the length
\item \textit{columns}: also orientation specific, for row-based blocks this is the length, for column-based blocks the size
\end{itemize}

@@

\end{table*}

\subsection{Constraint templates}
The goal is to learn constraints over blocks in the data. The knowledge needed to learn a constraint is expressed through {\template}s.
%
A \template $s$ is a triple $s=\textit{(\CName, \CSignature, \CFunction)}$:
%Let us elaborate on this:
\begin{itemize}
\item
\textit{\CName}  specifies the syntactic form of the constraint $s(B_1, ...,B_n)$, that is, the name of the template together
with $n$ abstract arguments $B_i$.
Thus a constraint is viewed as a relation or predicate of arity $n$ in first order logic.
Note that a function $B_r=f(B_1,...,B_n)$ can be represented with the $(n{+}1)$-ary predicate $s_f(B_r,B_1,...,B_n)$.
Each argument will have to be instantiated with a block.
% the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.

@@

  \item \CFunction: each value in the vector $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$.
\end{itemize}
\end{example}

@@

\newcommand{\sigc}{\ensuremath{\format{Sig}_s}}
\newcommand{\defc}{\ensuremath{\format{Def}_s}}
%\tias{TODO: conceptually $c(A_1, ..., A_n) :- Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$ where '$c$' is here both the constraint name and the template}.
With this interpretation, we can speak about the signature and definition of a constraint template being \textit{satisfied}.
\samuel{A: how do you get from the signature and the definition to these predicates? I guess this is what I'm missing on the formal side above… this needs to be made explicit}
We say that a signature (definition) of a constraint template~$s$ with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied.
Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming, we would write a Prolog-like clause: $s(B_1, ..., B_n) \leftarrow \sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
Under this interpretation, the term constraint and constraint template can be used interchangeably.

\begin{definition}
A \textbf{valid argument assignment} of a constraint $s$ is a tuple $(B_1, ..., B_n)$ such that $s(B_1, ..., B_n)$ is satisfied, that is, both the signature and the definition of the corresponding constraint template are satisfied by the assignment of $(B_1, ..., B_n)$ to the arguments.
\end{definition}

@@

We define the {\bf Tabular Constraint Learning Problem} as follows:
%
\samuel{A: so you expect the user to define blocks? can this be automated?
(also, later on, your solution seems to start from the tables rather than the blocks, why not match the two?)}
\begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
{\bf Given} a set of instantiated blocks ${\cal B}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $s(B'_1, ..., B'_n)$ where $s \in {\cal S}$ and $(B'_1, ..., B'_n)$ is a valid argument assignment of the template~$s$.
\end{definition}
% Let us now define what it means for a constraint from \template to hold in general. Let $t$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, and $C$ be the constraint $t^S$, then $C$ \textit{holds} iff all constraints in $\text{\CSignature}^S$ and $\text{\CFunction}^S$ of $C$ hold.
% \sergey{moved from formalization END}

@@

\subsubsection{Redundancies}
\label{sec:form:redundancies}
Depending on the application, some constraints in the solution to the tabular constraint learning problem may be considered \textit{redundant}. This is because constraints may be logically equivalent or may be implied by other constraints.

@@

Because dealing with redundancy is often application-dependent, we explain in the next section our generic method for finding all constraints. We describe some optimisations that avoid obvious equivalences in Section~\ref{sec:opts} and will investigate the impact of redundancy in the experiments.

@@

Our proposed methodology contains the following steps:
\begin{enumerate}
\item Extract headerless tables from tabular data
\item Partition the tables into maximally contiguous, type-consistent \textit{superblocks}
\item Generate for each constraint template all valid argument assignments in two steps:
\begin{enumerate}
\item For each constraint $c$, generate all atoms $c(B_1, \ldots ,B_n)$ where each $B_i$ is a maximal \textit{superblock} and the $B_i$ are compatible with the signature.
\item For each generated atom $c(B_1, \ldots ,B_n)$, find all valid $c(B'_1, \ldots, B'_n)$ such that for all $i$ holds $B'_i \sqsubseteq B_i$ and the signature and definition of $c(B'_1, \ldots, B'_n)$ are satisfied.
\end{enumerate}
\end{enumerate}

@@

Instead, we divide this step into two parts: in the first part, we will not reason over individual (sub)blocks and their data, but rather over the properties of the maximal superblocks.
% and what combination of superblocks may lead to a valid argument assignment.\\
Consider Table 1 and 2 in Figure~\ref{fig:main_example} and the $\mathit{SUM}_{col}$ constraint template.
Instead of considering all possible subblocks of T1 and T2, we first reason over the properties of the superblocks.
For example, $B2$ and $B4$ are not numeric so they can be immediately discarded.
\samuel{Iff Br is from T2?}
Looking at the sizes, $B1$ and~$B2$ contain only one and two columns respectively, however, the rows in $T2$ have length 4 so we only need to consider superblocks with at least 4 columns, which only $B3$ satisfies.

In the second part, we start from an atom $c(B_1, \ldots ,B_n)$ with maximal superblocks $B_i$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks. As we only have to consider the blocks $B'_i \sqsubseteq B_i$ contained in each superblock, this is typically much easier. For the above example, each of the vectors of $B6$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks for the right-hand side and verify the signature (e.g. at least size $4$) and test the definition for each of the rows. %In practice, this can be optimized further.
\\\\
\samuel{Section references not correct}
We now describe in more detail how the headerless tables are extracted and how the maximal superblocks are generated from them (step 1 and 2, Section 4.1), how the candidate superassignments are generated (step 3a, Section 4.2) and how the actual assignments are extracted from that (step 3b, Section 4.3). We then describe a number of optimizations that we perform to speed up step 3 and 4 (Section 4.4).

@@

This separation of checking the \textit{properties} of superblock assignments from checking the \textit{actual data} in the subblock assignment controls the exponential blow-up of combinations to test. Furthermore, we will use constraint satisfaction technology in the first step to efficiently enumerate superblock assignments that are compatible with the signature. %These are the two key properties of our approach.

@@

    \Procedure{LearnConstraints}{$\blocks$, $\temps$}
      \State $C \gets \emptyset$ %\Comment{The set of constraints}
      \ForAll{$s~\mathbf{in}~\temps$}
        %\State $n \gets$ number of arguments of template $s$
        \State $A \gets \generategroups(s, \blocks)$
        \ForAll{$(B_1, \dots, B_n) \in A$}
        \State $A' \gets \findassignment(s, (B_1,\dots,B_n))$
        \ForAll{$(B_1',\dots,B_n') \in A'$}
            \State $C \gets C \cup \{ c_s(B_1', \dots, B_n') \}$
          \EndFor
        \EndFor
      \EndFor
      \State \Return $C$
    \EndProcedure
\end{algorithmic}
\caption{Learn tabular constraints}
\label{algo:tcl}
\end{algorithm}

\subsubsection{Generating superblock assignments}
% Let there be $b$ blocks and let~$s$ have~$n$ arguments, then a simple generate-and-test method would have to test $b^n$ combinations (given that blocks can be repeated), which can grow very rapidly with $b$, depending on $n$.

@@

Table~\ref{tbl:translation} shows how the conversion from requirements of the signature to CSP constraints happens:
Requirements on the lengths, orientations and tables of blocks can be directly enforced since they are invariant under block containment($\sqsubseteq$).
Typing requirements need to be relaxed to check only for base types (numeric and textual).
Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since blocks with two many vectors contain subblocks of the required size.
Finally, restrictions on the number of rows or columns behave as length or size constraints based on the orientation of the block they are applied to.
Subgroups of row-oriented (column-oriented) blocks will always have the same number of columns (rows), however, the number of rows (columns) might decrease.

@@

\subsubsection{Generating subblock assignments}
\label{sec:algo:subgr}
Given a superblock assignment $(B_1, \dots, B_n)$ from the previous step, the goal is to discover valid \textit{subassignments}, i.e. assignments of subblocks $(B'_1, \dots, B'_n)$ (for all $i$, $B'_i \subseteq B_i$) that satisfy both the signature and the definition of template~$s$.

%While the previous step reasoned about the \textit{properties} of blocks, in this step we will also reason about the actual \textit{content} of the blocks, that is, the data.
\begin{example}
  Consider the sum-over-rows constraint template $\ecsumr{\sg_r}{\mathbf{\sg_x}}$.
  An example superblock assignment from Figure~\ref{fig:main_example} is $(B_r, \mathbf{\sg_x}) = (\range{T_1}{\rangeall}{\rangeto{3}{8}}, \range{T_1}{\rangeall}{\rangeto{3}{8}})$.
  Note that this assignment does not satisfy the signature yet, $\sg_r$ contains more than one vector and $\psize(\mathbf{\sg_x}) \neq \plength(B_r)$.
  This step aims to generate subassignments $(B_r' \sqsubseteq B_r, \mathbf{B_x}' \sqsubseteq \mathbf{B_x})$ that \textit{do} satisfy the signature and test whether they satisfy the definition: $\forall i, \sg_r[0][i] = \sum_j \mathbf{\sg_x}[j][i]$, where $B[i][j]$ indicates the $i$th value of the $j$th vector in $B$.
  In Figure~\ref{fig:main_example} there is exactly one such satisfying subassignment: $(T_1[:7], T_1[:,3:6])$.
\end{example}

In this step, we could also formulate the problem as a CSP. %However, the search space is much smaller, e.g. $O(nm^2)$ with $n$ the number of arguments of the template and $m$ the size of the largest block.
However, few CSP solvers support floating point numbers, which are prevalent in spreadsheets.
Furthermore, in a CSP approach we would have to \textit{ground} out the definition for each of the corresponding elements in the vectors.
\samuel{Tias, there is something wrong here:}
This is inefficient, as the signature of the definition may already not be satisfied, or the first element in the data.

Hence, a simple generate-and-test will usually suffice.
%However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
%For this reason, the \sname system uses generate-and-test to implement $\findassignment$ for most constraint templates.
Given a superblock assignment, all disjoint subassignments are generated taking into account the required size, columns or rows.
The disjoint subassignments are then sequentially tested using a template-specific procedure \textit{Test} and valid subassignments are returned.
The \textit{Test} procedure verifies the exact signature (e.g. subtypes will not have been checked yet) and checks whether the definition is satisfied for all data in the blocks.

\begin{algorithm}[tbh]
  \begin{algorithmic}
    \footnotesize
    \Procedure{Subassignments}{$s$, $(B_1, \dots, B_n)$}
    \State $n \gets$ number of arguments of template $s$
    \State $A_{sub} \gets \emptyset$
    \ForAll{$(B_1', \dots, B_n') \in \{ (B_1', \dots, B_n') | \forall i: B_i' \sqsubseteq B_i  \wedge \nexists i,j,i\neq j: B'_i \sqsubseteq B'_j \}$}
      \If{$Test((B_1',\dots,B_n'))$}
        \State $A_{sub} \gets A_{sub} \cup \{ (B_1', \dots, B_n') \}$
      \EndIf
    \EndFor
\Return $A_{sub}$
\EndProcedure
\end{algorithmic}
\caption{Generate-and-test for $\findassignment$}
\label{algo:gen-test-subblocks}
\end{algorithm}

@@

\begin{example}
  Consider $\ecfkey{B_{fk}}{B_{pk}}$ whose signature includes $\ecalldiff{B_{pk}}$ and assume that~$C$ are the constraints found for \textit{ALLDIFFERENT}.
  Instead of generating assignments for~$B_{pk}$ using the given blocks~$\blocks$, we will only consider (sub)blocks~$P = \{ B' | \ecalldiff{B'} \in C \}$.
  $P$~can be seen as a set of partial assignments in which the blocks for certain arguments have been fixed.
\end{example}

@@

As a result, in some cases a maximal subblock might falsely include irrelevant columns.
Consider $B_1 = [[200, 300], [200, 150], [1, 2]]$ and $B_2 = [200, 300]$, then \sname will find the constraint $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:3])}$.
Should, the target constraint be $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:2])}$, then the last vector of~$B_1$ must be split off into a separate block before the above approach can find it.

@@

The implementation is illustrated using a case study on the spreadsheet corresponding with the previously introduced example (figure~\ref{fig:main_example}).
In order to quantify the results and generalize our findings we also evaluate our algorithm on a benchmark of 30 (\samuel{check number}) spreadsheets that we assembled from various sources.

@@

\samuel{Revisit [what does "implemented" mean here? what's the difference between "implemented" and "algorithm", esp given that we are talking experiments? ]}
For every spreadsheet a ground-truth has been provided manually, specifying the essential (or original) constraints that are expected to be discovered.
We consider three types of constraints:
\begin{description}
  \item[Implemented] \samuel{Dupl} Constraint that have been implemented and are expected to be found (all constraints in Table~\ref{table:constraints})
  \item[Essential] \samuel{Dupl} Constraint that have or have not been implemented but could be found using our algorithm (e.g. fuzzy conditional sum)
  \item[Non-trivial] Constraints currently outside of the scope of this system (e.g. generic nested mathematical or logical formulas or n-ary constraints)
\end{description}

@@

Redundant constraints can be reduced in different ways.samuel
Duplications should be detected in a post-processing step that uses entailment or heuristics to filter constraints.
Accidental constraints may be detected through heuristics, however, they can be also be removed by adding additional data.

>>>>

@@

\title{Learning constraints in spreadsheets and tabular data}
% see http://www.acm.org/binaries/content/assets/publications/article-templates/sig-alternate-sample.tex

@@

\begin{abstract}
%Spreadsheets, comma separated value files and other tabular data representations are in wide use today. However, the correct use and maintenance of functions in and between the tabular data can be error-prone and overwhelming. In this work, we investigate the automatic discovery of constraints (functions and relations) from raw tabular data. Our method takes inspiration form inductive logic programming and constraint satisfiability. We represent common spreadsheet functions as predicates with arguments that must satisfy number of constraints over the arguments. This allows generic constraint satisfaction techniques to be used to find all such satisfying predicates/functions within the rows, columns and submatrices of the tables. We show the effectiveness and accuracy of this approach on a number of spreadsheets from varying sources.
Spreadsheets, CSV (comma separated value) files and other tabular data representations are in wide use today.
However, modeling, maintaining and discovering formulas in tabular data and spreadsheets can be time consuming and error-prone.
In this work, we investigate the automatic discovery of constraints (functions and relations) from raw tabular data.
We see multiple promising applications for this technique, e.g. in rediscovering constraints, auto-completion and error checking.
Our method takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
Common spreadsheet functions are represented as predicates whose arguments must satisfy a number of constraints.
Constraint learning techniques are used to identify predicates (constraints) that hold between blocks of rows and columns of the tables.
%New types of constraints can be easily added to the system by specifying them in a declarative language.
We show that our approach is able to accurately discover constraints in spreadsheets from various sources.
\end{abstract}

@@

The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where the discovery of clausal constraints has been studied in inductive logic programming \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.  It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from
very few examples.  Our approach contrasts with these in that it focusses
on learning both column- and row-constraints, as does Modelseeker~\cite{modelseeker} which is restricted to traditional constraint satisfaction problems (CSPs). It contrasts with Flashfill in that it can learn from numeric data too, as well as general constraints in addition to functions.
A more detailed discussion on related work is contained in Section~\ref{sec:related_work}.

@@

\section{Formalization}\label{sec:formalization}
Our goal is to automatically discover constraints (functions and relations) between the rows and the columns of tables in a spreadsheet. This is applicable not just to data from spreadsheets, but to any data in tabular form, hence the name.

@@

We first introduce some terminology and the concept of a \template, after which we define the problem and make some additional considerations.

@@

A block has the following properties:
\begin{itemize}
\item \textit{type}: a block is type-consistent, so it has a type
\item \textit{table}: the table that the block belongs to
\item \textit{orientation}: either row-oriented or column-oriented
\item \textit{size}: the number of vectors a block contains
\item \textit{length}: the length of its vectors; as all vectors are from the same table, they always have the same length.
\\\\
\luc{these are the vectors along \dots along}
\item \textit{rows}: this is orientation specific, for row-based blocks this is the size, for column-based blocks the length
\item \textit{columns}: also orientation specific, for row-based blocks this is the length, for column-based blocks the size
\end{itemize}

@@

  \luc{there are some fancy constraints here. it might be good to say something about them ... like an example of one of the more complex ones in the text ... maybe state what conditional means, what aggregate ... }
\end{table*}

\subsection{Constraint templates}
The goal is to learn constraints over blocks in the data. The knowledge needed to learn a constraint is expressed through {\template}s.
%
A \template $s$ is a triple $t = \textit{(\CName, \CSignature, \CFunction)}$:
%Let us elaborate on this:
\begin{itemize}
\item
\textit{\CName}  specifies the syntactic form of the constraint $s(B_1, ...,B_n)$, that is, the name of the template together
with $n$ abstract arguments $B_i$.
Thus a constraint is viewed as a relation or predicate of arity $n$ in first order logic.
Note that a function $B_r=f(B_1,...,B_n)$ can be represented with the $(n{+}1)$-ary predicate $s_f(B_r,B_1,...,B_n)$.
Each argument will have to be instantiated with a block.
% the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.

@@

  \item \CFunction: each value in the vector $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$. \luc{example where it holds}
\end{itemize}
\end{example}

@@

\newcommand{\sigc}{\ensuremath{\format{Sig}_s}}
\newcommand{\defc}{\ensuremath{\format{Def}_s}}
%\tias{TODO: conceptually $c(A_1, ..., A_n) :- Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$ where '$c$' is here both the constraint name and the template}.
With this interpretation, we can speak about the signature and definition of a constraint template being \textit{satisfied}.
\samuel{A: how do you get from the signature and the definition to these predicates? I guess this is what I'm missing on the formal side above… this needs to be made explicit}
We say that a signature (definition) of a constraint template~$s$ with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied.
Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming and logical relational learning, we would define the predicate~$s$ using a Prolog like clause: $s(B_1, ..., B_n) \leftarrow \sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
Under this interpretation, the term constraint and constraint template can be used interchangeably.

\begin{definition}
A \textbf{valid argument assignment} of a constraint $s$ is a tuple of blocks $(B_1, ..., B_n)$ such that $s(B_1, ..., B_n)$ is satisfied, that is, both the signature and the definition of the corresponding constraint template are satisfied by the assignment of $(B_1, ..., B_n)$ to the arguments.
\end{definition}

@@

We define the inverse problem, that is the {\bf Tabular Constraint Learning Problem}, as follows:
%
\samuel{A: so you expect the user to define blocks? can this be automated?
(also, later on, your solution seems to start from the tables rather than the blocks, why not match the two?)}
\begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
{\bf Given} a set of instantiated blocks ${\cal B}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $s(B'_1, ..., B'_n)$ where $s \in {\cal S}$, $\forall i: B_i' \sqsubseteq B_i \in {\cal B}$ and $(B'_1, ..., B'_n)$ is a satisfied argument assignment of the template~$s$.
\end{definition}
% Let us now define what it means for a constraint from \template to hold in general. Let $t$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, and $C$ be the constraint $t^S$, then $C$ \textit{holds} iff all constraints in $\text{\CSignature}^S$ and $\text{\CFunction}^S$ of $C$ hold.
% \sergey{moved from formalization END}

@@

\subsubsection{Redundancies}
\label{sec:form:redundancies}
Depending on the application, some constraints in the solution to the tabular constraint learning problem may be considered \textit{redundant}. This is because constraints may be logically equivalent or may be implied by other constraints.
\luc{if you have permutation, alldiff is redundant.}

@@

Dealing with redundancy is often application-dependent.
Therefore, in Section~\ref{sec:approach} we will first explain our generic method for finding all constraints, before describing some optimizations that avoid obvious equivalences (Section~\ref{sec:opts}).
Later, we will investigate the impact of redundancy in the experiments.

@@

Our proposed methodology contains the following steps:
\begin{enumerate}
\item Extract headerless tables from tabular data
\item Partition the tables into maximally contiguous, type-consistent \textit{superblocks}
\item Generate for each constraint template all valid argument assignments in two steps:
\begin{enumerate}
\item For each constraint $c$, generate all atoms $c(B_1, \ldots ,B_n)$ where each $B_i$ is a maximal superblock and the $B_i$ are \textit{compatible} (defined below) with the signature. 
\item For each generated atom $c(B_1, \ldots ,B_n)$ in 3a), find all valid $c(B'_1, \ldots, B'_n)$ such that for all $i$ holds $B'_i \sqsubseteq B_i$ and the signature and definition of $c(B'_1, \ldots, B'_n)$ are satisfied.
\end{enumerate}
\end{enumerate}
\luc{introduce the notion of a maximal superblock earlier ???}

@@

Instead, we divide this step into two parts: in step 3a, we will not reason over individual (sub)blocks and their data, but rather over the properties of the maximal superblocks.
% and what combination of superblocks may lead to a valid argument assignment.\\
Consider tables $T_1$ and $T_2$ in Figure~\ref{fig:main_example} and the $\mathit{SUM}_{col}$ constraint template.
Instead of considering all possible subblocks of $T_1$ and $T_2$, we first reason over the properties of the superblocks.
For example, $B_2$ and $B_4$ are not numeric so they can be immediately discarded.
\samuel{Iff Br is from T2?}
Looking at the sizes, $B_1$ and~$B_2$ contain only one and two columns respectively, however, the rows in $T_2$ have length 4 so we only need to consider superblocks with at least 4 columns, which only $B_3$ satisfies. \luc{so the generated atom in step 3a is \dots }

In step 3b, we start from an atom $c(B_1, \ldots, B_n)$ with maximal superblocks $B_i$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks. As we only have to consider the blocks $B'_i \sqsubseteq B_i$ contained in each superblock, this is typically much easier. For the above example, each of the vectors of $B_6$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks for the right-hand side and verify the signature (e.g. at least size $4$) and test whether for each row the definition is satisfied. %In practice, this can be optimized further.
\\\\
\samuel{Section references not correct}
We now describe in more detail how the headerless tables are extracted and how the maximal superblocks are generated from them (step 1 and 2, Section 4.1), how the candidate superassignments are generated (step 3a, Section 4.2) and how the actual assignments are extracted from that (step 3b, Section 4.3). We then describe a number of optimizations that we perform to speed up steps 3 and 4 (Section 4.4).

@@

This separation of checking the \textit{properties} of superblock assignments from checking the \textit{actual data} in the subblock assignment controls the exponential blow-up of combinations to test. Furthermore, we use constraint satisfaction technology in the first step to efficiently enumerate superblock assignments that are compatible with the signature. %These are the two key properties of our approach.

@@

    \Procedure{LearnConstraints}{$\blocks$, $\temps$}
      \State $C \gets \emptyset$ %\Comment{The set of constraints}
      \ForAll{$s~\mathbf{in}~\temps$}
        %\State $n \gets$ number of arguments of template $s$
        \State $A \gets \generategroups(s, \blocks)$
        \ForAll{$(B_1, \dots, B_n) \in A$}
        \State $A' \gets \findassignment(s, (B_1,\dots,B_n))$
        \ForAll{$(B_1',\dots,B_n') \in A'$}
            \Comment \luc{if cs(B; …( satisfied/?}
            \State $C \gets C \cup \{ c_s(B_1', \dots, B_n') \}$
          \EndFor
        \EndFor
      \EndFor
      \State \Return $C$
    \EndProcedure
\end{algorithmic}
\caption{Learn tabular constraints}
\label{algo:tcl}
\end{algorithm}

\subsubsection{Generating superblock assignments (Step 3a)}
% Let there be $b$ blocks and let~$s$ have~$n$ arguments, then a simple generate-and-test method would have to test $b^n$ combinations (given that blocks can be repeated), which can grow very rapidly with $b$, depending on $n$.

@@

Table~\ref{tbl:translation} shows how the conversion from requirements of the signature to CSP constraints works:
Requirements on the lengths, orientations and tables of blocks can be directly enforced since they are invariant under block containment($\sqsubseteq$).
Typing requirements need to be relaxed to check only for base types (numeric and textual).
Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since blocks with two many vectors contain subblocks of the required size.
Finally, restrictions on the number of rows or columns behave as length or size constraints based on the orientation of the block they are applied to.
Subgroups of row-oriented (column-oriented) blocks will always have the same number of columns (rows), however, the number of rows (columns) might decrease.

@@

\subsubsection{Generating subblock assignments (Step 3b)}
\label{sec:algo:subgr}
Given a superblock assignment $(B_1, \dots, B_n)$ from the previous step, the goal is to discover valid \textit{subassignments}, i.e. assignments of subblocks $(B'_1, \dots, B'_n)$ (for all $i$, $B'_i \subseteq B_i$) that satisfy both the signature and the definition of template~$s$.

%While the previous step reasoned about the \textit{properties} of blocks, in this step we will also reason about the actual \textit{content} of the blocks, that is, the data.
\begin{example}
  Consider the sum-over-rows constraint template $\ecsumr{\sg_r}{\mathbf{\sg_x}}$.
  An example superblock assignment from Figure~\ref{fig:main_example} is $(B_r, \mathbf{\sg_x}) = (\range{T_1}{\rangeall}{\rangeto{3}{8}}, \range{T_1}{\rangeall}{\rangeto{3}{8}})$.
  Note that this assignment does not satisfy the signature yet, $\sg_r$ contains more than one vector and $\psize(\mathbf{\sg_x}) \neq \plength(B_r)$.
  This step aims to generate subassignments $(B_r' \sqsubseteq B_r, \mathbf{B_x}' \sqsubseteq \mathbf{B_x})$ that \textit{do} satisfy the signature and test whether they satisfy the definition: $\forall i, \sg_r[0][i] = \sum_j \mathbf{\sg_x}[j][i]$, where $B[i][j]$ indicates the $i$th value of the $j$th vector in $B$.
  In Figure~\ref{fig:main_example} there is exactly one such satisfying subassignment: $(T_1[:7], T_1[:,3:6])$. \luc{why [j][i] and not [j,i] ?}
\end{example}

In this step, we could also formulate the problem as a CSP. %However, the search space is much smaller, e.g. $O(nm^2)$ with $n$ the number of arguments of the template and $m$ the size of the largest block.
However, few CSP solvers support floating point numbers, which are prevalent in spreadsheets.
Furthermore, in a CSP approach we would have to \textit{ground} out the definition for each of the corresponding elements in the vectors.
\samuel{Tias, there is something wrong here:}
This is inefficient, as already the signature of the definition may not be satisfied, or the first element in the data.

Hence, a simple generate-and-test will usually suffice.
%However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
%For this reason, the \sname system uses generate-and-test to implement $\findassignment$ for most constraint templates.
\luc{test = definition, why use another term ?}
Given a superblock assignment, all disjoint subassignments are generated taking into account the required size, columns or rows.
The disjoint subassignments are then sequentially tested using a template-specific procedure \textit{Test} and valid subassignments are returned.
The \textit{Test} procedure verifies the exact signature (e.g. subtypes will not have been checked yet) and checks whether the definition is satisfied for all data in the blocks.

\begin{algorithm}[tbh]
  \begin{algorithmic}
    \footnotesize
    \Procedure{Subassignments}{$s$, $(B_1, \dots, B_n)$}
    \State $n \gets$ number of arguments of template $s$
    \State $A_{sub} \gets \emptyset$
    \ForAll{$(B_1', \dots, B_n') \textbf{ where } \forall i: B_i' \sqsubseteq B_i  \land \textit{disjoint}_{j \neq i}(B_i', B_j')$}
      \If{$Test((B_1',\dots,B_n'))$}
      \Comment \luc{Test = def is satisfied}
        \State $A_{sub} \gets A_{sub} \cup \{ (B_1', \dots, B_n') \}$
      \EndIf
    \EndFor
\Return $A_{sub}$
\EndProcedure
\end{algorithmic}
\caption{Generate-and-test for $\findassignment$}
\label{algo:gen-test-subblocks}
\end{algorithm}

@@

\begin{example}
  Consider $\ecfkey{B_{fk}}{B_{pk}}$ whose signature includes $\ecalldiff{B_{pk}}$ and assume that~$C$ are the constraints found for \textit{ALLDIFFERENT}.
  Instead of generating assignments for~$B_{pk}$ using the given blocks~$\blocks$, we will only consider (sub)blocks~$P = \{ B' | \ecalldiff{B'} \in C \}$.
  $P$~can be seen as a set of partial assignments in which the argument $B_{pk}$ has been fixed. \samuel{is it clear now?}
\end{example}

@@

As a result, in some cases a maximal subblock might falsely include irrelevant columns.
Consider $B_1 = [[200, 300], [200, 150], [1, 2]]$ and $B_2 = [200, 300]$, then \sname will find the constraint $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:3])}$.
Should the target constraint be $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:2])}$, then the last vector of~$B_1$ must be split off into a separate block before the above approach can find it.

@@

The implementation is illustrated using a case study on the spreadsheet corresponding to the previously introduced example (Figure~\ref{fig:main_example}).
In order to quantify the results and generalize our findings we also evaluate our algorithm on a benchmark of 30 (\samuel{check number}) spreadsheets that we assembled from various sources.

@@

\samuel{Revisit [what does "implemented" mean here? what's the difference between "implemented" and "algorithm", esp given that we are talking experiments? ] }
\luc{what is not implemented, strange \dots in a paper !!!}
For every spreadsheet a ground-truth has been provided manually, specifying the essential (or original) constraints that are expected to be discovered.
We consider three types of constraints:
\begin{description}
  \item[Implemented] \samuel{Dupl} Constraint that have been implemented and are expected to be found (all constraints in Table~\ref{table:constraints})
  \item[Essential] \samuel{Dupl} Constraint that have or have not been implemented but could be found using our algorithm (e.g. fuzzy conditional sum)
  \item[Non-trivial] Constraints currently outside of the scope of this system (e.g. generic nested mathematical or logical formulas or n-ary constraints)
\end{description}

@@

Redundant constraints can be reduced in different ways.
Duplications should be identified in a post-processing step that uses entailment or heuristics to filter constraints.
Accidental constraints may be detected through heuristics, however, they can be also be removed by adding additional data.
